Synthesizing Quotient Lenses

### 2017-10-17 papers/QuotientLensSynth/main.tex
Comments and a bit of working on the abstract (needs more)

  {\em Quotient lenses} express bidirectional transformations between data
  formats, where the inputs to the transformation are considered modulo some
  equivalence relation.  Such lenses save the programmer from having to write
  two expressions and satisfy strong ``lens laws'' that ensure that the forward
  and backward transformations interact well.  The use of quotients allows the
  programmer to ignore ``inessential'' differences in the data, such as the
  amount whitespace or the ordering of fields.

  However, quotient lenses are often difficult to program.  The programmer
  often has to give a specification for the data on each side of the lens and a
  separate specification for the equivalence relations defined on the data. The
  programmer must also manually implement a lens program that maps between the
  specifications while respecting the equivalence relations---this is itself a
  challenging task thanks to the restrictions on how lenses can be combined
  while still satisfying the lens laws.

This paper addresses both of these challenges. First, we propose the
language of {\em Quotient Regular Expressions}, or QREs, which 
define a broad class of equivalence relations
directly on regular languages using a compact and convenient notation. This
eliminates the need to specify the data and the equivalence relation on the data
separately. We then define {\em QRE lenses}, a class of quotient
lenses that are typed using QREs, and show how to {\em synthesize} QRE lenses
from their QRE types and a set of input-output examples. Finally, we
demonstrate the applicability of this approach by synthesizing QRE lenses
between several real-world data formats from the {\tt data.gov} database.
\end{abstract}

@@

Bidirectional programs have also been called {\em lenses}. Formally, Foster et
al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
following equations:
\begin{align*}
\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
v \; s) &= v\\
\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
\end{align*}

@@

\bcp{I thought we were going to allow newlines within titles and authors...?}

\sam{I wanted to illustrate how to use the squash QRE which is why I changed my
mind on that, but on second thought since the problem of quotienting away
whitespace at will is going to be more familiar to the reader I changed it back
to the original plan. I will explain squash in the QRE section.}
\begin{verbatim}
@Book {conway,
  Author = "J. H. Conway",
  Title = {Regular Algebra and Finite Machines},
}

@@

The code sample below shows how the programmer might specify these equivalence
relations on the \bibtex{} records in Boomerang using canonizers:

@@

With QREs we can specify \bibtex{} records as follows:

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{attheedges}
  \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
  screenshot from Nate's thesis. Are we allowed to use it?}}
  \label{fig:attheedges}
\end{figure}

@@

>>>>

{\em Lenses}~\cite{Lenses} are programs that specify bidirectional
transformations between different data formats.  Writing a single lens
rather than two separate uni-directional transformations saves both
keystrokes and bugs, by reducing redundancy and by automatically
guaranteeing strong ``lens laws'' ensuring that the forward and backward
transformations interact well.
%
{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
where the two data formats are considered modulo some equivalence relation.
The use of quotients allows the programmer to ignore inessential differences
in the data, such as the amount of whitespace or the ordering of fields.

Quotient lenses can be difficult to program, requiring the programmer to
give both a specification for the data on each side of the lens and a
separate specification for the equivalence relations on the data.  The
programmer must also manually implement a lens program that maps between the
formats while respecting the equivalence relations---this is itself a
challenging task thanks to the restrictions on how lenses can be combined
while still satisfying the lens laws.  \bcp{Don't really understand that
  paragraph!}

This paper addresses both of these challenges\bcp{what are the two
  challenges?}. First, we propose the language of {\em Quotient Regular
  Expressions}, or QREs, which define a broad class of equivalence relations
directly on regular languages using a compact and convenient notation. This
eliminates the need to specify the data and the equivalence relation on the
data separately. We then define {\em QRE lenses}, a class of quotient lenses
that are typed using QREs, and show how to {\em synthesize} QRE lenses from
their QRE types and a set of input-output examples. Finally, we demonstrate
the applicability of this approach by synthesizing QRE lenses between
several real-world data formats from the {\tt data.gov} database.
\end{abstract}

@@

Bidirectional programs have also been called {\em lenses}. Formally, Foster et
al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
following equations:\bcp{Why are we showing all the complexity of asymmetric
lenses?  We should just say that we're considering only the bijective case
for now and then just give the details of that.}
\begin{align*}
\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
v \; s) &= v\\
\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
\end{align*}

@@

\begin{verbatim}
@Book {conway,
  Author = "J. H. Conway",
  Title = {Regular Algebra and Finite Machines},
}

@@

The code sample below shows how the programmer might specify these equivalence
relations on the \bibtex{} records in Boomerang using canonizers:\bcp{Too
  big to display in-line.  Better to put it in a figure.}

@@

With QREs we can specify \bibtex{} records as follows:\bcp{Again, belongs in
a figure.  And, since both figures are long, maybe there is a visual way to
highlight the differences...?}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{attheedges}
  \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
      screenshot from Nate's thesis. Are we allowed to use it?}\bcp{We can
      (if we put ``from [Nate's thesis]'' in the caption), but I don't think
      we want the full complexity of asymmetric lenses here.  We should draw
      the figure again.  (I might have the OmniGraffle file for it lying
      around someplace...)}}
  \label{fig:attheedges}
\end{figure}

@@

\bcp{No bibliography??}
